---
title: "1 — Pre-Processing: GLCM Textures & Vegetation Indices"
author: 
  name: "Kyle Cahitas"
  email: "kcahitas@calpoly.edu"
date: today
format: html
editor: source
execute:
  error: true
  echo: true
  message: false
  warning: false
---
# Notes Before Reading
Hi Ian,
Below is the workflow the NR-418 group used. 
A separate qmd file will be created for the Sequoia Riverlands Trust. You can run all this code but it may be insufficient time. We included this workflow to show you our logic. Thank you!

# Libraries

```{r}
library(terra)
library(sf)
library(tidyverse)
library(viridis)
library(GLCMTextures)
library(RColorBrewer)
library(tidyterra)

# Custom color palettes for vegetation
veg_colors <- colorRampPalette(c("brown", "yellow", "lightgreen", "darkgreen"))
```

# Loading the Original Raster

```{r}
raster_200 <- rast("og_tif_files/og_merge.tif")
```

# Getting Bands

```{r}
red_200 <- raster_200[[1]]
green_200 <- raster_200[[2]]  
blue_200 <- raster_200[[3]]
```

# Running GLCM

There was an error with this code, had to run ONE band each, export the raster and then clear the environment and restart.

Don't Run — Already ran but good code to have.

Want to unhash it all: on a Mac use Cmd + Shift + C

## Green GLCM

```{r}
# textr_green <- glcm_textures(
#   green_200,
#   w           = c(7, 7),
#   n_levels    = 16,
#   quant_method = "prob",
#   shift       = list(c(1, 0),
#                      c(1, 1),
#                      c(0, 1),
#                      c(-1, 1)),
#   metrics     = c("glcm_contrast",
#                   "glcm_homogeneity",
#                   "glcm_entropy",
#                   "glcm_variance",
#                   "glcm_dissimilarity",
#                   "glcm_mean",
#                   "glcm_correlation"))
# 
# plot(textr_green)
```

### Green GLCM export

```{r}
# writeRaster(textr_green, "full_textures/glcm_200_green_all_bands.tif", overwrite = TRUE)
```

```{r}
# # Check layer names
# names(textr_green)
# 
# # Split and save each texture metric as its own .tif
# for (i in 1:nlyr(textr_green)) {
#   layer_name <- names(textr_green)[i]
#   writeRaster(
#     textr_green[[i]],
#     filename  = paste0("full_textures/GLCM_Green", layer_name, ".tif"),
#     overwrite = TRUE
#   )
# }
```

## Red GLCM

```{r}
# textr_red <- glcm_textures(
#   red_200,
#   w           = c(7, 7),          
#   n_levels    = 16,              
#   quant_method = "prob",          
#   shift       = list(c(1, 0),    
#                      c(1, 1),
#                      c(0, 1),
#                      c(-1, 1)),
#   metrics     = c("glcm_contrast",
#                   "glcm_homogeneity",
#                   "glcm_entropy",
#                   "glcm_variance",
#                   "glcm_dissimilarity",
#                   "glcm_mean",
#                   "glcm_correlation"))
# 
# plot(textr_red)
```

### Red GLCM export

```{r}
# writeRaster(textr_red, "full_textures/glcm_200_red_all_bands.tif", overwrite = TRUE)
```

```{r}
# names(textr_red)
# 
# # Split and save each texture metric as its own .tif
# for (i in 1:nlyr(textr_red)) {
#   layer_name <- names(textr_red)[i]
#   writeRaster(
#     textr_red[[i]],
#     filename  = paste0("full_textures/GLCM_Red", layer_name, ".tif"),
#     overwrite = TRUE
#   )
# }
```

## Blue GLCM

```{r}
# textr_blue <- glcm_textures(
#   blue_200,
#   w           = c(7, 7),         
#   n_levels    = 16,              
#   quant_method = "prob",          
#   shift       = list(c(1, 0),    
#                      c(1, 1),
#                      c(0, 1),
#                      c(-1, 1)),
#   metrics     = c("glcm_contrast",
#                   "glcm_homogeneity",
#                   "glcm_entropy",
#                   "glcm_variance",
#                   "glcm_dissimilarity",
#                   "glcm_mean",
#                   "glcm_correlation"))
# 
# plot(textr_blue)
```

## Blue GLCM export

```{r}
# writeRaster(textr_blue, "full_textures/glcm_200_blue_all_bands.tif", overwrite = TRUE)
```

```{r}
# # Check layer names
# names(textr_blue)
# 
# # Split and save each texture metric as its own .tif
# for (i in 1:nlyr(textr_blue)) {
#   layer_name <- names(textr_blue)[i]
#   writeRaster(
#     textr_blue[[i]],
#     filename  = paste0("full_textures/GLCM_Blue", layer_name, ".tif"),
#     overwrite = TRUE
#   )
# }
```

# Other Indices

Again, these will all be in a folder, don't run.

```{r}
red <- red_200 
green <- green_200 
blue <- blue_200 
```

## VARI (Visible Atmospherically Resistant Index)

```{r}
# **Formula:** VARI = (Green - Red) / (Green + Red - Blue)
vari <- (green - red) / (green + red - blue)
names(vari) <- "VARI"

# Handle division by zero or very small denominators
vari <- ifel((green + red - blue) == 0, NA, vari)

# Visualize
plot(vari, main="VARI (Visible Atmospherically Resistant Index)",
     col=veg_colors(100))

writeRaster(vari, "full_textures/VARI_200.tif", overwrite = TRUE)
```

## GLI (Green Leaf Index)

```{r}
# **Formula:** GLI = (2×Green - Red - Blue) / (2×Green + Red + Blue)
gli <- (2*green - red - blue) / (2*green + red + blue)
names(gli) <- "GLI"

# Handle division by zero
gli <- ifel((2*green + red + blue) == 0, NA, gli)

# Visualize
plot(gli, main="GLI (Green Leaf Index)",
     col=veg_colors(100))
writeRaster(gli, "full_textures/GLI_200.tif", overwrite = TRUE)
```

## ExG (Excess Green Index)

```{r}
# **Formula:** ExG = 2×Green - Red - Blue
exg <- 2*green - red - blue
names(exg) <- "ExG"

# Normalize to 0-1 range for visualization
exg_norm <- (exg - global(exg, min, na.rm=TRUE)[1,1]) / 
            (global(exg, max, na.rm=TRUE)[1,1] - global(exg, min, na.rm=TRUE)[1,1])

# Visualize
par(mfrow=c(1,2))
plot(exg, main="ExG (Excess Green)", col=veg_colors(100))
plot(exg_norm, main="ExG (Normalized)", col=veg_colors(100))
par(mfrow=c(1,1))
writeRaster(exg, "full_textures/EXG_200.tif", overwrite = TRUE)
writeRaster(exg_norm, "full_textures/EXG_norm_200.tif", overwrite = TRUE)
```

## NGRDI (Normalized Green-Red Difference Index)

```{r}
# **Formula:** NGRDI = (Green - Red) / (Green + Red)
ngrdi <- (green - red) / (green + red)
names(ngrdi) <- "NGRDI"

# Handle division by zero
ngrdi <- ifel((green + red) == 0, NA, ngrdi)

# Visualize
plot(ngrdi, main="NGRDI (Normalized Green-Red Difference Index)",
     col=veg_colors(100))
writeRaster(ngrdi, "full_textures/NGRDI_200.tif", overwrite = TRUE)
```

## TGI (Triangular Greenness Index)

```{r}
# **Formula:** TGI = Green - 0.39×Red - 0.61×Blue
tgi <- green - 0.39*red - 0.61*blue
names(tgi) <- "TGI"

# Visualize
plot(tgi, main="TGI (Triangular Greenness Index)",
     col=veg_colors(100))
writeRaster(tgi, "full_textures/TGI_200.tif", overwrite = TRUE)
```

## RGB Vegetation Index (RGBVI)

```{r}
rgbvi <- (green^2 - red * blue) / (green^2 + red * blue)
names(rgbvi) <- "RGBVI"

# Handle division by zero
rgbvi <- ifel((green^2 + red * blue) == 0, NA, rgbvi)

# Visualize
plot(rgbvi, main="Red Green Blue Vegetation Index (RGBVI)",
     col=veg_colors(100))
writeRaster(rgbvi, "full_textures/RGBVI_200.tif", overwrite = TRUE)
```

## ExGR

```{r}
exr <- (1.4 * red) - green
names(exr) <- "ExR"
# Formula: ExGR = ExG - ExR
exgr <- exg - exr
names(exgr) <- "ExGR"
exgr <- ifel(is.infinite(exgr), NA, exgr)
plot(exgr, main="Excess Green Red Index (ExGR)",
     col=veg_colors(100))
writeRaster(exgr, "full_textures/EXGR_200.tif", overwrite = TRUE)
```

## Vegetation Index (VEG)

```{r}
# Formula: VEG = Green / (Red^0.667 * Blue^0.333)
veg <- green / (red^0.667 * blue^0.333)
names(veg) <- "VEG"
veg <- ifel((red^0.667 * blue^0.333) == 0, NA, veg)
veg <- ifel(is.infinite(veg), NA, veg)
plot(veg, main="Vegetation Index (VEG)",
     col=veg_colors(100))
writeRaster(veg, "full_textures/VEG_200.tif", overwrite = TRUE)
```
