---
title: "2 — PCA: Principal Component Analysis on Raster Stack"
author: 
  name: "Kyle Cahitas"
  email: "kcahitas@calpoly.edu"
date: today
format: html
editor: source
execute:
  error: true
  echo: true
  message: false
  warning: false
---
# Notes Before Reading
Hi Ian,
Below is the workflow the NR-418 group used. 
A separate qmd file will be created for the Sequoia Riverlands Trust. You can run all this code but it may be insufficient time. We included this workflow to show you our logic. Thank you!

# Libraries

```{r}
library(terra)
library(sf)
library(tidyverse)
library(viridis)
library(GLCMTextures)
library(RColorBrewer)
library(factoextra)
library(corrplot)

# Custom color palettes for vegetation
veg_colors <- colorRampPalette(c("brown", "yellow", "lightgreen", "darkgreen"))
```

# PCA

## Importing the files

This loop loads all .tif files from the full_textures folder.

```{r}
data_folder <- "full_textures"
file_list <- list.files(path = data_folder, pattern = "\\.tif{1,2}$", full.names = TRUE)

for (file in file_list) {
  og_name <- tools::file_path_sans_ext(basename(file))
  assign(og_name, rast(file))
  cat("Loaded:", og_name, "\n")
}
```

There were a lot of errors, but this was the only way to correctly see the tif files in the PCA results.

```{r}
# Red GLCM
glcm_red_contrast     <- GLCM_Redglcm_contrast
glcm_red_correlation  <- GLCM_Redglcm_correlation
glcm_red_dissimilarity<- GLCM_Redglcm_dissimilarity
glcm_red_entropy      <- GLCM_Redglcm_entropy
glcm_red_homogeneity  <- GLCM_Redglcm_homogeneity
glcm_red_mean         <- GLCM_Redglcm_mean
glcm_red_variance     <- GLCM_Redglcm_variance

# Green GLCM
glcm_green_contrast      <- GLCM_Greenglcm_contrast
glcm_green_correlation   <- GLCM_Greenglcm_correlation
glcm_green_dissimilarity <- GLCM_Greenglcm_dissimilarity
glcm_green_entropy       <- GLCM_Greenglcm_entropy
glcm_green_homogeneity   <- GLCM_Greenglcm_homogeneity
glcm_green_mean          <- GLCM_Greenglcm_mean
glcm_green_variance      <- GLCM_Greenglcm_variance

# Blue GLCM
glcm_blue_contrast      <- GLCM_Blueglcm_contrast
glcm_blue_correlation   <- GLCM_Blueglcm_correlation
glcm_blue_dissimilarity <- GLCM_Blueglcm_dissimilarity
glcm_blue_entropy       <- GLCM_Blueglcm_entropy
glcm_blue_homogeneity   <- GLCM_Blueglcm_homogeneity
glcm_blue_mean          <- GLCM_Blueglcm_mean
glcm_blue_variance      <- GLCM_Blueglcm_variance

# RGB Indices
ExG_rgb   <- EXG_200
ExGR_rgb  <- EXGR_200
RGBVI_rgb <- RGBVI_200
TGI_rgb   <- TGI_200
VARI_rgb  <- VARI_200
VEG_rgb   <- VEG_200
NGRDI_rgb <- NGRDI_200
GLI_rgb   <- GLI_200
```

```{r}
# Red GLCM
names(glcm_red_contrast)      <- "red_contrast"
names(glcm_red_correlation)   <- "red_correlation"
names(glcm_red_dissimilarity) <- "red_dissimilarity"
names(glcm_red_entropy)       <- "red_entropy"
names(glcm_red_homogeneity)   <- "red_homogeneity"
names(glcm_red_mean)          <- "red_mean"
names(glcm_red_variance)      <- "red_variance"

# Green GLCM
names(glcm_green_contrast)      <- "green_contrast"
names(glcm_green_correlation)   <- "green_correlation"
names(glcm_green_dissimilarity) <- "green_dissimilarity"
names(glcm_green_entropy)       <- "green_entropy"
names(glcm_green_homogeneity)   <- "green_homogeneity"
names(glcm_green_mean)          <- "green_mean"
names(glcm_green_variance)      <- "green_variance"

# Blue GLCM
names(glcm_blue_contrast)      <- "blue_contrast"
names(glcm_blue_correlation)   <- "blue_correlation"
names(glcm_blue_dissimilarity) <- "blue_dissimilarity"
names(glcm_blue_entropy)       <- "blue_entropy"
names(glcm_blue_homogeneity)   <- "blue_homogeneity"
names(glcm_blue_mean)          <- "blue_mean"
names(glcm_blue_variance)      <- "blue_variance"

# RGB Indices
names(ExG_rgb)   <- "ExG"
names(ExGR_rgb)  <- "ExGR"
names(RGBVI_rgb) <- "RGBVI"
names(TGI_rgb)   <- "TGI"
names(VARI_rgb)  <- "VARI"
names(VEG_rgb)   <- "VEG"
names(NGRDI_rgb) <- "NGRDI"
names(GLI_rgb)   <- "GLI"
```

```{r}
raster_pca <- c(glcm_red_contrast, glcm_red_correlation, glcm_red_dissimilarity,
         glcm_red_entropy, glcm_red_homogeneity, glcm_red_mean, glcm_red_variance,
         glcm_green_contrast, glcm_green_correlation, glcm_green_dissimilarity,
         glcm_green_entropy, glcm_green_homogeneity, glcm_green_mean, glcm_green_variance,
         glcm_blue_contrast, glcm_blue_correlation, glcm_blue_dissimilarity,
         glcm_blue_entropy, glcm_blue_homogeneity, glcm_blue_mean, glcm_blue_variance,
         ExG_rgb, ExGR_rgb, RGBVI_rgb, TGI_rgb, VARI_rgb, VEG_rgb, NGRDI_rgb, GLI_rgb)

print(names(raster_pca))

raster_stack <- raster_pca
```

# Summary Table

```{r}
summary_stats <- terra::global(raster_stack, fun = c("mean", "sd"), na.rm = TRUE)
print(summary_stats)
```

Dropping ExGR due to high levels.

```{r}
raster_stack <- c(glcm_red_contrast, glcm_red_correlation, glcm_red_dissimilarity,
         glcm_red_entropy, glcm_red_homogeneity, glcm_red_mean, glcm_red_variance,
         glcm_green_contrast, glcm_green_correlation, glcm_green_dissimilarity,
         glcm_green_entropy, glcm_green_homogeneity, glcm_green_mean, glcm_green_variance,
         glcm_blue_contrast, glcm_blue_correlation, glcm_blue_dissimilarity,
         glcm_blue_entropy, glcm_blue_homogeneity, glcm_blue_mean, glcm_blue_variance,
         ExG_rgb, RGBVI_rgb, TGI_rgb, VARI_rgb, VEG_rgb, NGRDI_rgb, GLI_rgb)
```

## Running PCA now

```{r}
set.seed(805)
raster_sample <- terra::spatSample(raster_stack, size = 10000,
                                method = "random", na.rm = TRUE, as.df = TRUE)
raster_sample <- na.omit(raster_sample)

pca_result <- prcomp(raster_sample, center = TRUE, scale. = TRUE)

var_exp   <- pca_result$sdev^2
var_pct   <- var_exp / sum(var_exp) * 100
cum_pct   <- cumsum(var_pct)

var_table <- tibble(
  PC         = paste0("PC", seq_along(var_pct)),
  Eigenvalue = round(var_exp, 3),
  Variance   = round(var_pct, 2),
  Cumulative = round(cum_pct, 2)
)

print(var_table, n = 30)
```

Already ran this code — outputs are in pca_results/.

```{r}
# pc_rasters <- terra::predict(raster_stack, pca_result, index = 1:6)
# 
# # Save as a single multi-band GeoTIFF
# writeRaster(pc_rasters,
#             filename  = "pca_results/PCA_PC1_to_PC6.tif",
#             overwrite = TRUE)
# names(pc_rasters)
# 
# # Split and save each texture metric as its own .tif
# for (i in 1:nlyr(pc_rasters)) {
#   layer_name <- names(pc_rasters)[i]
#   writeRaster(
#     pc_rasters[[i]],
#     filename  = paste0("pca_results/PCA", layer_name, ".tif"),
#     overwrite = TRUE
#   )
# }
```
