---
title: "3 — Teacher Model: Full-Feature Random Forest Classifier"
author: 
  name: "Kyle Cahitas"
  email: "kcahitas@calpoly.edu"
date: today
format: html
editor: source
execute:
  error: true
  echo: true
  message: false
  warning: false
---
# Notes Before Reading
Hi Ian,
Below is the workflow the NR-418 group used. 
A separate qmd file will be created for the Sequoia Riverlands Trust. You can run all this code but it may be insufficient time. We included this workflow to show you our logic. Thank you!

# Libraries

```{r}
library(terra)
library(sf)
library(tidyverse)
library(viridis)
library(GLCMTextures)
library(RColorBrewer)
library(factoextra)
library(corrplot)
library(randomForest)
library(caret)
library(ranger)
library(e1071)
library(MLmetrics)
library(doParallel)
```

I ran code for all the layers, just in case the global environment was wrong. A bit redundant but helps visually see everything again.

## Importing the files

```{r}
pca_1 <- rast("pca_results/pca_1.tif")
pca_2 <- rast("pca_results/pca_2.tif")
pca_3 <- rast("pca_results/pca_3.tif")
pca_4 <- rast("pca_results/pca_4.tif")
pca_5 <- rast("pca_results/pca_5.tif")
pca_6 <- rast("pca_results/pca_6.tif")
```

```{r}
data_folder <- "full_textures"
file_list <- list.files(path = data_folder, pattern = "\\.tif{1,2}$", full.names = TRUE)

for (file in file_list) {
  og_name <- tools::file_path_sans_ext(basename(file))
  assign(og_name, rast(file))
  cat("Loaded:", og_name, "\n")
}
```

```{r}
# Red GLCM
glcm_red_contrast     <- GLCM_Redglcm_contrast
glcm_red_correlation  <- GLCM_Redglcm_correlation
glcm_red_dissimilarity<- GLCM_Redglcm_dissimilarity
glcm_red_entropy      <- GLCM_Redglcm_entropy
glcm_red_homogeneity  <- GLCM_Redglcm_homogeneity
glcm_red_mean         <- GLCM_Redglcm_mean
glcm_red_variance     <- GLCM_Redglcm_variance

# Green GLCM
glcm_green_contrast      <- GLCM_Greenglcm_contrast
glcm_green_correlation   <- GLCM_Greenglcm_correlation
glcm_green_dissimilarity <- GLCM_Greenglcm_dissimilarity
glcm_green_entropy       <- GLCM_Greenglcm_entropy
glcm_green_homogeneity   <- GLCM_Greenglcm_homogeneity
glcm_green_mean          <- GLCM_Greenglcm_mean
glcm_green_variance      <- GLCM_Greenglcm_variance

# Blue GLCM
glcm_blue_contrast      <- GLCM_Blueglcm_contrast
glcm_blue_correlation   <- GLCM_Blueglcm_correlation
glcm_blue_dissimilarity <- GLCM_Blueglcm_dissimilarity
glcm_blue_entropy       <- GLCM_Blueglcm_entropy
glcm_blue_homogeneity   <- GLCM_Blueglcm_homogeneity
glcm_blue_mean          <- GLCM_Blueglcm_mean
glcm_blue_variance      <- GLCM_Blueglcm_variance

# RGB Indices
ExG_rgb   <- EXG_200
ExGR_rgb  <- EXGR_200
RGBVI_rgb <- RGBVI_200
TGI_rgb   <- TGI_200
VARI_rgb  <- VARI_200
VEG_rgb   <- VEG_200
NGRDI_rgb <- NGRDI_200
GLI_rgb   <- GLI_200

# PCA layers
pca_1 <- pca_1
pca_2 <- pca_2
pca_3 <- pca_3
pca_4 <- pca_4
pca_5 <- pca_5
pca_6 <- pca_6
```

```{r}
# Red GLCM
names(glcm_red_contrast)      <- "red_contrast"
names(glcm_red_correlation)   <- "red_correlation"
names(glcm_red_dissimilarity) <- "red_dissimilarity"
names(glcm_red_entropy)       <- "red_entropy"
names(glcm_red_homogeneity)   <- "red_homogeneity"
names(glcm_red_mean)          <- "red_mean"
names(glcm_red_variance)      <- "red_variance"

# Green GLCM
names(glcm_green_contrast)      <- "green_contrast"
names(glcm_green_correlation)   <- "green_correlation"
names(glcm_green_dissimilarity) <- "green_dissimilarity"
names(glcm_green_entropy)       <- "green_entropy"
names(glcm_green_homogeneity)   <- "green_homogeneity"
names(glcm_green_mean)          <- "green_mean"
names(glcm_green_variance)      <- "green_variance"

# Blue GLCM
names(glcm_blue_contrast)      <- "blue_contrast"
names(glcm_blue_correlation)   <- "blue_correlation"
names(glcm_blue_dissimilarity) <- "blue_dissimilarity"
names(glcm_blue_entropy)       <- "blue_entropy"
names(glcm_blue_homogeneity)   <- "blue_homogeneity"
names(glcm_blue_mean)          <- "blue_mean"
names(glcm_blue_variance)      <- "blue_variance"

# RGB Indices 
names(ExG_rgb)   <- "ExG"
names(ExGR_rgb)  <- "ExGR"
names(RGBVI_rgb) <- "RGBVI"
names(TGI_rgb)   <- "TGI"
names(VARI_rgb)  <- "VARI"
names(VEG_rgb)   <- "VEG"
names(NGRDI_rgb) <- "NGRDI"
names(GLI_rgb)   <- "GLI"

# PCA layers
names(pca_1) <- "pca_1"
names(pca_2) <- "pca_2"
names(pca_3) <- "pca_3"
names(pca_4) <- "pca_4"
names(pca_5) <- "pca_5"
names(pca_6) <- "pca_6"
```

```{r}
raster_stack <- c(glcm_red_contrast, glcm_red_correlation, glcm_red_dissimilarity,
         glcm_red_entropy, glcm_red_homogeneity, glcm_red_mean, glcm_red_variance,
         glcm_green_contrast, glcm_green_correlation, glcm_green_dissimilarity,
         glcm_green_entropy, glcm_green_homogeneity, glcm_green_mean, glcm_green_variance,
         glcm_blue_contrast, glcm_blue_correlation, glcm_blue_dissimilarity,
         glcm_blue_entropy, glcm_blue_homogeneity, glcm_blue_mean, glcm_blue_variance,
         ExG_rgb, ExGR_rgb, RGBVI_rgb, TGI_rgb, VARI_rgb, VEG_rgb, NGRDI_rgb, GLI_rgb,
         pca_1, pca_2, pca_3, pca_4, pca_5, pca_6)

print(names(raster_stack))  # Double Check
```

Remember that we used the entire drone raster; code below will now split it into two. The right side is the test and the left side is the validation.

```{r}
# DON'T RUN (if you want to) — will provide all of the files
# e <- ext(raster_stack)
# mid_x <- (e$xmin + e$xmax) / 2
# 
# full_vald <- ext(e$xmin, mid_x, e$ymin, e$ymax) |> 
#   crop(x = raster_stack)
# 
# writeRaster(full_vald, "output_tifs/full_vald.tif", overwrite = TRUE)
# 
# full_test <- ext(mid_x, e$xmax, e$ymin, e$ymax) |> 
#   crop(x = raster_stack)
# 
# writeRaster(full_test, "full_textures/full_test.tif", overwrite = TRUE)
```

```{r}
full_test <- rast("full_textures/full_test.tif")
```

# Stacking the polygons now

## Importing polygons

```{r}
bg_poly  <- vect("polygons/BG_poly_test/BG_poly_test.shp")
ds_poly  <- vect("polygons/DS_poly_test/DS_poly_test.shp")
grs_poly <- vect("polygons/GRS_poly_test/GRS_poly_test.shp")
rt_poly  <- vect("polygons/RT_poly_test/RT_poly_test.shp")
sb_poly  <- vect("polygons/SB_poly_test/SB_poly_test.shp")
shd_poly <- vect("polygons/SHD_poly_test/SHD_poly_test.shp")
```

## Beginning Preprocessing steps

```{r}
bg_vals <- terra::extract(full_test, bg_poly, cells = TRUE) |> 
  mutate(class = "bare_ground")

ds_vals <- terra::extract(full_test, ds_poly, cells = TRUE) |> 
  mutate(class = "dead_shrubs")

grs_vals <- terra::extract(full_test, grs_poly, cells = TRUE) |> 
  mutate(class = "grass")

rt_vals <- terra::extract(full_test, rt_poly, cells = TRUE) |> 
  mutate(class = "russian_thistle")

sb_vals <- terra::extract(full_test, sb_poly, cells = TRUE) |> 
  mutate(class = "saltbush")

shd_vals <- terra::extract(full_test, shd_poly, cells = TRUE) |> 
  mutate(class = "shadow")
```

## Stacking them all

```{r}
train_data <- bind_rows(
  bg_vals,
  ds_vals,
  grs_vals,
  rt_vals,
  sb_vals,
  shd_vals) |> 
  select(-ID) |>
  na.omit() |>
  mutate(class = factor(class, levels = c("saltbush", "russian_thistle", "shadow", "grass", "dead_shrubs", "bare_ground"))) |> 
  arrange(class) |>
  filter(!duplicated(cell)) |>
  dplyr::select(-any_of(c("ID", "id", "cell"))) |>
  mutate(class = as.factor(class))

cat("Raw class distribution:\n")
```

```{r}
# Save the train data frame, just in case
saveRDS(train_data, "rds_outputs/train_data.rds")
```

## Train Data Visual

```{r}
train_data |>
  count(class) |>
  ggplot(aes(x = class, 
             y = n, 
             fill = class)) +
  geom_col() +
  geom_text(aes(label = n), 
            vjust = -0.5) +
  scale_fill_brewer(palette = "YlGnBu") +
  labs(title = "Class Distribution (Before Balancing)", 
       y = "Pixel Count") +
  theme_minimal()
```

## Balanced Train Data Visual

```{r}
min_count <- train_data |> 
  count(class) |> 
  pull(n) |> 
  min()

set.seed(42)
balanced_data <- train_data |>
  group_by(class) |>
  slice_sample(n = min_count) |>
  ungroup()

balanced_data |> 
  count(class) |> 
  print()

balanced_data |>
  count(class) |>
  ggplot(aes(x = class, 
             y = n, 
             fill = class)) +
  geom_col() +
  geom_text(aes(label = n), vjust = -0.5) +
  scale_fill_brewer(palette = "YlGnBu") +
  labs(title = "Class Distribution (After Balancing)", 
       y = "Pixel Count") +
  theme_minimal()
```

```{r}
set.seed(805)

train_set <- balanced_data |> 
  slice_sample(prop = 0.7)

test_set  <- balanced_data |> 
  anti_join(train_set)

cat("\nTraining:", nrow(train_set), "| Testing:", nrow(test_set), "\n")
```

# Caret (Machine Learning — Teacher Model)

Note: We originally called it "mother model" but changed it to "teacher model" for the presentation. mother model = teacher model; child model = student model.

```{r}
# Load saved train/test sets if resuming from a previous session
# train_set <- read_rds("rds_outputs/train_set.rds")
# test_set  <- read_rds("rds_outputs/test_set.rds")

# For reference: original machine is MacBook Pro M4 48GB — adjust cores if needed
cl <- makePSOCKcluster(parallel::detectCores() - 2)
registerDoParallel(cl)

train_control <- trainControl(
  method          = "repeatedcv",
  number          = 10,
  repeats         = 3,
  classProbs      = TRUE,
  summaryFunction = multiClassSummary,
  savePredictions = "final",
  allowParallel   = TRUE)

tune_grid <- expand.grid(
  mtry          = c(2, 4, 6, 8),
  splitrule     = "extratrees",
  min.node.size = 1)

set.seed(123)
rf_caret <- train(
  class ~ .,
  data       = train_set,
  method     = "ranger",
  trControl  = train_control,
  tuneGrid   = tune_grid,
  importance = "permutation",
  num.trees  = 500)

stopCluster(cl)
registerDoSEQ()

write_rds(rf_caret, "rds_outputs/rf_caret_model.rds")
write_rds(train_set, "rds_outputs/train_set.rds")
write_rds(test_set, "rds_outputs/test_set.rds")
print(rf_caret)
```

# Identifying labels now to transfer for the student model

```{r}
mother_model <- readRDS("rds_outputs/rf_caret_model.rds")
```

```{r}
mother_stack <- rast("full_textures/full_test.tif")
```

```{r}
# Chunk-based prediction to avoid memory issues
n_chunks <- 10
rows_per_chunk <- ceiling(nrow(mother_stack) / n_chunks)

# Create chunk extents
chunks <- lapply(1:n_chunks, function(i) {
  row_start <- (i - 1) * rows_per_chunk + 1
  row_end <- min(i * rows_per_chunk, nrow(mother_stack))
  ext(mother_stack, r1 = row_start, r2 = row_end)
})

# Predict each chunk sequentially but save to disk
chunk_files <- c()
for (i in 1:n_chunks) {
  cat("Processing chunk", i, "of", n_chunks, "\n")
  
  chunk_rast <- crop(mother_stack, chunks[[i]])
  
  chunk_file <- paste0("chunk_", i, ".tif")
  chunk_files <- c(chunk_files, chunk_file)
  
  terra::predict(
    chunk_rast, mother_model,
    type = "raw",
    filename = chunk_file,
    overwrite = TRUE)
  
  # Free memory
  rm(chunk_rast)
  gc()
}
```

```{r}
# Merge all chunks back together
cat("Merging chunks...\n")
chunk_rasts <- lapply(chunk_files, rast)
mother_labels <- do.call(merge, chunk_rasts)

writeRaster(mother_labels, "output_tifs/mother_labels.tif", overwrite = TRUE)
names(mother_labels) <- "class"

file.remove(chunk_files)
```

Quick visual

```{r}
mother_labels <- rast("output_tifs/mother_labels.tif")
```

```{r}
plot(mother_labels, type = "classes",
     col = c("#2D6A4F", "#E6393D", "#333333", "#95D5B2", "#BC6C25", "#DDB892"),
     main = "Mother Model Labels")
```
