---
title: "Final Template"
author:
  name: "Kyle Cahitas"
  email: "kcahitas@calpoly.edu"
date: today
format: html
editor: source
execute:
  echo: true
  eval: false
---

# Notes Before Reading
For Sequoia Riverlands Trust. This code and model only work for drone RGB imagery. Every code chunk has a `#| eval: false`. If you want to use this template, you need to delete that.
 
# Installation
Unhash and run once, if you dont have these packages. 
```{r}
#| eval: false
#install.packages("terra")
#install.packages("sf")
#install.packages("tidyverse")
#install.packages("caret")
#install.packages("GLCMTextures")
#install.packages("ranger")
#install.packages("e1071")
#install.packages("RColorBrewer")
#install.packages("doParallel")
```

# Libraires
```{r}
#| eval: false
library(terra)
library(sf)
library(tidyverse)
library(caret)
library(ranger)
library(GLCMTextures)
library(e1071)
library(RColorBrewer)
library(doParallel)
```

# Loading Data
Loading Carrizo Plain Drone Imagery. Ensure that the Drone Imagery is in EPSG:3857 - WGS 84 / Pseudo-Mercator.

```{r}
#| eval: false
drone_raster <- rast("DRONE_IMAGERY.tif") #change to the name of your file & correlate it with the data file, I have set up 
child_model <- readRDS("rds_outputs/child_model.rds")
red <- drone_raster[[1]]
green <- drone_raster[[2]]
blue <- drone_raster[[3]]
```

# Note
Remember, for the model to work, the new drone imagery needs to have the exact same bands, and the GLCM textures need to be computed in the same way.

## GLCM Textures 
Note. There could be a better streamlined version, but I separate each band and then run the GLCM Texture function.
```{r}
#| eval: false
# Green 
textr_green <- glcm_textures(
  green,
  w           = c(7, 7),
  n_levels    = 16,
  quant_method = "prob",
  shift       = list(c(1, 0),
                     c(1, 1),
                     c(0, 1),
                     c(-1, 1)),
  metrics     = c("glcm_contrast",
                  "glcm_homogeneity",
                  "glcm_entropy",
                  "glcm_variance"))
writeRaster(textr_green, "glcm_200_green_all_bands.tif", overwrite = TRUE) #this is just insurance, take awhile to run 

# Red
textr_red <- glcm_textures(
  red,
  w           = c(7, 7),
  n_levels    = 16,
  quant_method = "prob",
  shift       = list(c(1, 0),
                     c(1, 1),
                     c(0, 1),
                     c(-1, 1)),
  metrics     = c("glcm_contrast",
                  "glcm_homogeneity",
                  "glcm_entropy",
                  "glcm_variance"))
writeRaster(textr_red, "glcm_200_red_all_bands.tif", overwrite = TRUE) #this is just insurance, take awhile to run 

# Blue
textr_blue <- glcm_textures(
  blue,
  w           = c(7, 7),
  n_levels    = 16,
  quant_method = "prob",
  shift       = list(c(1, 0),
                     c(1, 1),
                     c(0, 1),
                     c(-1, 1)),
  metrics     = c("glcm_contrast",
                  "glcm_homogeneity",
                  "glcm_entropy",
                  "glcm_variance"))
writeRaster(textr_blue, "glcm_200_blue_all_bands.tif", overwrite = TRUE) #this is just insurance, take awhile to run 
```

## Understanding the outputs of the glcm_texture() function

### Window function
```{r}
#| eval: false
glcm_textures(
  data, 
  w = c(7,7),
  ..., )# means there are more functions to connect to
```

This function examines each pixel using a 7 × 7 block of surrounding pixels to compute texture statistics.

### Levels function
```{r}
#| eval: false
glcm_textures(
  data, 
  ...,# means there are more functions to connect to
  n_levels = 16)
```
The raster pixels ranged from 0 to 255. Running the level functions bins down into 16 gray levels for the GLCM matrix. Originally had it at 32, but kept on crashing, so switched to 16.

### Shifts function
```{r}
#| eval: false
glcm_textures(
  data, 
  ..., # means there are more functions to connect to
  shift = list(c(1, 0),
                     c(1, 1),
                     c(0, 1),
                     c(-1, 1)))
```
Measures four directions:horizontal, diagonal down-right, vertical and diagonal down-left.


### Metrics function
```{r}
#| eval: false
glcm_textures(
  data, 
  ..., # means there are more functions to connect to
  metrics     = c("glcm_contrast",
                  "glcm_homogeneity",
                  "glcm_entropy",
                  "glcm_variance"))
```
The outputs you want. In this case its only 4 types of GLCM textures. 

## Vegetation Indices
```{r}
#| eval: false
tgi  <- green - 0.39 * red - 0.61 * blue
names(tgi) <- "TGI"
exg  <- 2 * green - red - blue
names(exg) <- "ExG"
vari <- (green - red) / (green + red - blue)
vari <- ifel((green + red - blue) == 0, NA, vari)
names(vari) <- "VARI"
```
Used only three Indice to simplify the model. These where the most important. 


# Renaming before stacking
There is probably a better way. The only reason we are renaming is to better understand the output.
```{r}
#| eval: false
# Red GLCM — extract layers from textr_red
glcm_red_contrast     <- textr_red[["glcm_contrast"]]
glcm_red_homogeneity  <- textr_red[["glcm_homogeneity"]]
glcm_red_entropy      <- textr_red[["glcm_entropy"]]
glcm_red_variance     <- textr_red[["glcm_variance"]]
##
names(glcm_red_contrast)      <- "red_contrast"
names(glcm_red_entropy)       <- "red_entropy"
names(glcm_red_homogeneity)   <- "red_homogeneity"
names(glcm_red_variance)      <- "red_variance"

# Green GLCM — extract layers from textr_green
glcm_green_contrast     <- textr_green[["glcm_contrast"]]
glcm_green_homogeneity  <- textr_green[["glcm_homogeneity"]]
glcm_green_entropy      <- textr_green[["glcm_entropy"]]
glcm_green_variance     <- textr_green[["glcm_variance"]]
##
names(glcm_green_contrast)      <- "green_contrast"
names(glcm_green_entropy)       <- "green_entropy"
names(glcm_green_homogeneity)   <- "green_homogeneity"
names(glcm_green_variance)      <- "green_variance"

# Blue GLCM — extract layers from textr_blue
glcm_blue_contrast     <- textr_blue[["glcm_contrast"]]
glcm_blue_homogeneity  <- textr_blue[["glcm_homogeneity"]]
glcm_blue_entropy      <- textr_blue[["glcm_entropy"]]
glcm_blue_variance     <- textr_blue[["glcm_variance"]]
##
names(glcm_blue_contrast)      <- "blue_contrast"
names(glcm_blue_entropy)       <- "blue_entropy"
names(glcm_blue_homogeneity)   <- "blue_homogeneity"
names(glcm_blue_variance)      <- "blue_variance"
```

```{r}
#| eval: false
names(red)   <- "red_b"
names(blue)  <- "blue_b"
names(green) <- "green_b"
```

# Stacking 
The names must also be the same for the model to work.
```{r}
#| eval: false
drone_stack <- c(glcm_blue_entropy, glcm_green_entropy, glcm_red_entropy,
                 glcm_green_contrast, glcm_blue_contrast, glcm_red_contrast,
                 glcm_blue_variance, glcm_green_variance, glcm_red_variance,
                 glcm_blue_homogeneity, glcm_green_homogeneity, glcm_red_homogeneity,
                 tgi, exg, vari, red, blue, green)
nlyr(drone_stack)
#check stack, make sure there is 18
```

# Running the child/student model and new raster stack
This model is taking the information from the child model and now applying it to the new drone stack. It classifies every pixel in the raster into one of six classes.

## Refresher of classes
```{r}
#"saltbush", "russian_thistle", "shadow","grass","dead_shrubs","bare_ground"
```

## Running code now
Combining the new drone data with the child/student model results.
```{r}
#| eval: false
n_chunks <- 10  # Increase to 20-30 if you have <16 GB RAM
rows_per_chunk <- ceiling(nrow(drone_stack) / n_chunks)
chunk_files <- c()
 
for (i in 1:n_chunks) {
  cat("Processing chunk", i, "of", n_chunks, "\n")
  row_start <- (i - 1) * rows_per_chunk + 1
  row_end   <- min(i * rows_per_chunk, nrow(drone_stack))
  full_ext  <- ext(drone_stack)
  res_y     <- yres(drone_stack)
  y_max <- full_ext$ymax - (row_start - 1) * res_y
  y_min <- full_ext$ymax - row_end * res_y
  chunk_ext  <- ext(full_ext$xmin, full_ext$xmax, y_min, y_max)
  chunk_rast <- crop(drone_stack, chunk_ext)
  chunk_file <- paste0("chunk_", i, ".tif")
  chunk_files <- c(chunk_files, chunk_file)
  terra::predict(
    chunk_rast, child_model,
    type = "raw", na.rm = TRUE,
    filename = chunk_file, overwrite = TRUE
  )
  rm(chunk_rast); gc()
}
# Merge all chunks
chunk_rasts <- lapply(chunk_files, rast)
classified  <- do.call(merge, chunk_rasts)
names(classified) <- "class"
writeRaster(classified, "my_classification.tif", overwrite = TRUE)

```
This method helps the computer run more efficiently. The function `n_chunk()` splits the data into 10 to help separate the large dataset. The `cat()` function helps show the progress of the functions because they take a while. The row numbers are converted to geographic coordinates and used to calculate pixel coordinates. The function `terra::predict()` runs the classification model on selecting each pixel. (Every time you run `predict()`, make sure to specify it with terra::, there is another package that uses that function name. The functions `rm()` and `gc()` help with the chunk memory and RAM.


# Classification Map 
```{r}
#| eval: false
class_colors <- c(
  "saltbush"        = "#2d6a4f",
  "russian_thistle" = "#e63946",
  "shadow"          = "#333333",
  "grass"           = "#95d5b2",
  "dead_shrubs"     = "#bc6c25",
  "bare_ground"     = "#ddb892"
)
 
plot(
  classified,
  main = "Land Cover Classification",
  col  = class_colors,
  mar  = c(3, 3, 3, 10),
  plg  = list(title = "Land Cover", cex = 0.9))
```

## Visual
```{r}
#| eval: false
freq_table <- freq(classified)
freq_table$percent <- round(
  freq_table$count / sum(freq_table$count) * 100, 2
)
print(freq_table)
```

# Confidence map 
It's called "confidence," but it is just the model's probability for each class. Ranging from 0-1, the closer to 1, the more "confident" the model is.

```{r}
#| eval: false
probabilities <- terra::predict(
  drone_stack, child_model,
  type     = "prob",
  na.rm    = TRUE,
  filename = "my_probabilities.tif",
  overwrite = TRUE)
#Extracting the Confidence Layer
probabilities <- rast("my_probabilities.tif")
confidence <- max(probabilities)
names(confidence) <- "confidence"
 
plot(confidence, main = "Classification Confidence",
     col = viridis::viridis(100))
 
writeRaster(confidence, "my_confidence.tif",
            overwrite = TRUE)
```

To check this model's accuracy on the drone raster stack, follow the steps for the teacher/mother model. Selecting polygons and seeing if the model is accurate. Or, you can import the classified raster stack and check visually. Stack both the original drone imagery and the classification model, toggle the transparency, and see if the model was correct.

# Sources that helped: 
1. https://rstudio-pubs-static.s3.amazonaws.com/253860_05f11cddd938407a9cb3b06d9dc38c9a.html 
2. https://www.wvview.org/os_sa/15b_Raster_Analysis_terra.html
3. https://github.com/rspatial/terra/issues/968
4. https://rspatial.github.io/terra/reference/predict.html 
5. https://rdrr.io/cran/caret/man/predict.train.html


# Questions
If you have any questions, please contact me at my email:kcahitas@calpoly.edu, and we can set up a Zoom meeting.




